Parsing Tweets into Universal Dependencies

### 2017-12-15 paper.tex
Noah

\begin{document}
\maketitle
\begin{abstract}
We study the problem of analyzing tweets with
universal dependencies \citep[UD;][]{NIVRE16.348}. We extend the UD guidelines to cover
special constructions in tweets that affect tokenization,
part-of-speech tagging, and dependencies. Using the extended guidelines, we create
a new tweet treebank ({\sc Tweebank v2}) that is four times larger than the (unlabeled) {\sc Tweebank
  v1} introduced by \citet{kong-EtAl:2014:EMNLP2014}. 
We characterize the disagreements between our annotators
and show that it is challenging to deliver
consistent annotation due to ambiguity in
understanding and explaining tweets. Nonetheless, using the new treebank,
we build a pipeline system to parse raw tweets into UD. To overcome the
annotation noise without sacrificing computational efficiency, we propose a new
method to distill an ensemble of 20 transition-based parsers into a single one. Our
parser achieves an improvement of 2.6 in LAS over the un-ensembled baseline 
and outperforms parsers that are state-of-the-art on other treebanks in both accuracy and speed.
\end{abstract}

@@

\item We introduce and evaluate a pipeline system to parse the raw tweet text into
universal dependencies (\S\ref{sec:parsing}).  Experimental results show
it performs better than a pipeline of the state-of-the-art alternatives.

@@

\begin{figure*}[t]
	\centering
	\small
	\begin{dependency}[edge slant=2, text only label, label style=above]
		\begin{deptext}
			Perfect \& \textcolor{red}{\bf $\heart$} \& \textcolor{orange}{\bf RT} \& \textcolor{orange}{\bf @coldplay} \& : \& Fix \& You \& from \& the \& back \& \textcolor{gray}{\bf \#ColdplayMinneapolis} \& \textcolor{blue}{\bf http://bit.ly/2dj2WCl} \& Nice \& \textcolor{cyan}{\bf p} \& ...\\
			\tiny ADJ \&\tiny SYM \&\tiny X \&\tiny X \&\tiny PUNCT \&\tiny VERB \&\tiny NOUN \&\tiny ADP \&\tiny DET \&\tiny NOUN \&\tiny X \&\tiny X \& \tiny ADJ \& \tiny X \& \tiny PUNCT \\
		\end{deptext}
		\deproot{1}{root}
		\depedge[edge unit distance=1em]{1}{2}{discourse}
		\deproot{6}{root}
		\depedge[edge unit distance=1em]{6}{3}{discourse}
		\depedge[edge unit distance=1em]{3}{4}{discourse}
		\depedge[edge unit distance=1em]{3}{5}{punct}
		\depedge[edge unit distance=1em]{6}{7}{obj}
		\depedge[edge unit distance=0.8em]{6}{10}{obl}
		\depedge[edge unit distance=1em]{10}{8}{case}
		\depedge[edge unit distance=1em]{10}{9}{det}
		\depedge[edge unit distance=0.8em]{6}{11}{discourse}
		\depedge[edge unit distance=0.8em]{6}{12}{list}
		\deproot{13}{root}		
		\depedge[edge unit distance=0.8em]{13}{14}{discourse}
		\depedge[edge unit distance=0.8em]{13}{15}{punct}
	\end{dependency}
	\caption{An example tweet that contains non-syntactic tokens:
		\textcolor{red}{sentiment emoticon},
		\textcolor{orange}{retweet marker and its following at-mention},
		\textcolor{gray}{topical hashtag},
		\textcolor{blue}{referential URL}, and
		\textcolor{cyan}{truncated word}.
		This example tweet is a concatenation of three real tweets.
	}\label{fig:non-syn-toks}
\end{figure*}

\begin{figure*}[t]
	\centering
	\small
	\begin{dependency}[edge slant=2, text only label, label style=above]
		\begin{deptext}
			@username \& its \& \#awesome \& u \& gonna \& $\heart$ \& it \& Chk \& out \& our \& cooool \& project \& on \& http://project\_link \& + \& RT \& it\\
			\tiny PROPN \& \tiny PRON \& \tiny ADJ \& \tiny PRON \& \tiny VERB \& \tiny VERB \& \tiny PRON \& \tiny VERB \& \tiny ADP \& \tiny PRON \& \tiny ADJ \& \tiny NOUN \& \tiny ADP \& \tiny X \& \tiny SYM \& \tiny VERB\& \tiny PRON\\
		\end{deptext}
		\deproot{3}{root}
		\depedge[edge unit distance=1em]{3}{1}{vocative}
		\depedge[edge unit distance=1em]{3}{2}{nsubj}
		\depedge[edge unit distance=1em]{3}{5}{parataxis}
		\depedge[edge unit distance=1em]{5}{4}{nsubj}
		\depedge[edge unit distance=1em]{5}{6}{xcomp}
		\depedge[edge unit distance=1em]{6}{7}{obj}
		\depedge[edge unit distance=0.7em]{3}{8}{parataxis}
		\depedge[edge unit distance=1em]{8}{9}{compound:prt}
		\depedge[edge unit distance=0.7em]{8}{12}{obj}
		\depedge[edge unit distance=1em]{12}{10}{nmod:poss}
		\depedge[edge unit distance=1em]{12}{11}{amod}
		\depedge[edge unit distance=0.6em]{8}{14}{obl}
		\depedge[edge unit distance=1em]{14}{13}{case}
		\depedge[edge unit distance=0.55em]{8}{16}{conj}
		\depedge[edge unit distance=1em]{16}{15}{cc}
		\depedge[edge unit distance=1em]{16}{17}{obj}
	\end{dependency}
	\caption{An example tweet with informal but syntactic tokens.
		This is a contrived example inspired by several tweets.}\label{fig:informal-toks}
\end{figure*}

@@

\begin{itemize}
%\item For the at-mentions in retweet constructions, we attach them to their retweet markers. We will discuss our choice in the following paragraph;
\item If a non-syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate. Retweet construction is a special case and we will discuss our choice in the following paragraph;
\item If the whole sentence is made of a sequence of non-syntactic tokens, we attach all these tokens to the first one;
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item Non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English dataset.% \yicomment{need to add punct for punctuations?}
%\\ \yicomment{For retweet at-mention case, we attach the retweet at-mentions to the RT sign, so that the retweet action and the people whose tweet is being retweeted could be connected.\footnote{We can also follow the first convention to attach the retweet at-mention to its predicate, but we think it is better for downstream tasks.}}
\end{itemize}

@@

\paragraph{Retweet construction.} Figure \ref{fig:non-syn-toks} shows
an example of the retweet construction (\emph{RT @coldplay :}).  This
might be treated as a verb phrase, with \emph{RT} as a verb and the
at-mention as its object\yicomment{@coldplay seems not to be an object, could be some modifier like vocative, or some prepositional case (the tweets on/about coldplay)}\yjcomment{i am okey with current version}.
% \yicomment{or might be treated as 2 non-syntactic tokens and a punctuation, which should all attach to the predicate of the following retweeted tweet.}
%\yjcomment{i wouldn't like the readers to think in this direction.}.
  This solution would lead to an uninformative root
word and, since this expression is idiomatic to Twitter, might create
unnecessary confusion for downstream applications aiming to identify
the main predicate(s) of a tweet.  We therefore treat the whole
expression as non-syntactic, including assigning the \emph{other}
(\texttt{X}) part of speech to both \emph{RT} and \emph{@coldplay},
attaching the at-mention to \emph{RT} with the \emph{discourse}
label and the colon to \emph{RT} with the \emph{punct}(uation) label,
and attaching \emph{RT} to the predicate of the following sentence.
%\yicomment{since we have introduced retweet at-mentions in the non-syntactic tokens, here we can just briefly justify our choice.}.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & $F_1$ \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\		 
% 		UD pipe v1.2 & 94.7 \\
		\hdashline
		UD pipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}
\citet{wang-EtAl:2017:Long6} shows that the parser trained 
on the combination of UD\_English and English dialect (Singlish) treebank
outperfrom the one only on the dialect treebank.
We followed their work and trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English training set
and tested on the \textsc{Tweebank v2} test set. \yicomment{should mention that combination of corpora improves the results, otherwise we should answer why we didn't try training ONLY on Tweebank v2}\yjcomment{add discussion}
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540} systems are rule-based systems and were not adapted
to the UD tokenization scheme.
The UD pipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. Compared with the UD pipe, we use LSTM
instead of GRU in our model and we also use a larger size of hidden units (64 against 20),
which has stronger representation power. 
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results indicate the value of statistical modeling
in tokenization for informal texts.

@@

Part-of-speech tagging for tweets has been extensively studied \cite{ritter-EtAl:2011:EMNLP,gimpel-EtAl:2011:ACL-HLT2011,derczynski-EtAl:2013:RANLP-2013,owoputi-EtAl:2013:NAACL-HLT,gui-EtAl:2017:EMNLP20172}. We therefore consider existing POS taggers for tweets instead of developing our own.
On the annotation scheme designed in \S\ref{sec:pos-anno}, based on UD and adapted for
Twitter, we compared several existing systems: the Stanford CoreNLP tagger, 
\citet{owoputi-EtAl:2013:NAACL-HLT}'s word cluster enhanced tagger
(both greedy and CRF variants), and 
\citet{ma-hovy:2016:P16-1}'s neural network tagger 
which achieves the state-of-the-art performance on PTB.
\citet{gui-EtAl:2017:EMNLP20172} presented the state-of-the-art neural tagger for Twitter,
but their code didn't work beyond PTB tag set. Thus, we didn't include their tagger in our comparison.\yjcomment{i put some words on \cite{gui-EtAl:2017:EMNLP20172} in case the reviewer challenge our neural/non-neural comparsion.}
 %and also two bi-LSTM systems similar
%to \citet{DBLP:journals/corr/HuangXY15} and \citet{lample-EtAl:2016:N16-1}, one using
%word vectors and the other using word and character vectors (see Figure \ref{fig:pos-model}).
All compared systems were re-trained on the combination of the UD\_English and 
\textsc{Tweebank v2} training sets. We use Twitter-specific GloVe embeddings released by
\citet{pennington-socher-manning:2014:EMNLP2014} in all neural taggers
and parsers.\footnote{\url{http://nlp.stanford.edu/data/glove.twitter.27B.zip}}

>>>>

\begin{document}
\maketitle
\begin{abstract}
We study the problem of analyzing tweets with
universal dependencies \citep[UD;][]{NIVRE16.348}. We extend the UD guidelines to cover
special constructions in tweets that affect tokenization,
part-of-speech tagging, and labeled dependencies. Using the extended guidelines, we create
a new tweet treebank ({\sc Tweebank v2}) that is four times larger than the (unlabeled) {\sc Tweebank
  v1} introduced by \citet{kong-EtAl:2014:EMNLP2014}. 
We characterize the disagreements between our annotators
and show that it is challenging to deliver
consistent annotation due to ambiguity in
understanding and explaining tweets. Nonetheless, using the new treebank,
we build a pipeline system to parse raw tweets into UD. To overcome the
annotation noise without sacrificing computational efficiency, we propose a new
method to distill an ensemble of 20 transition-based parsers into a single one. Our
parser achieves an improvement of 2.6 in LAS over the un-ensembled baseline 
and outperforms parsers that are state-of-the-art on other treebanks in both accuracy and speed.
\end{abstract}

@@

\item We introduce and evaluate a pipeline system to parse the raw tweet text into
universal dependencies (\S\ref{sec:parsing}).  Experimental results show
that it performs better than a pipeline of the state-of-the-art alternatives.

@@

\begin{figure*}[t]
	\centering
	\small
	\begin{dependency}[edge slant=2, text only label, label style=above]
		\begin{deptext}
			Perfect \& \textcolor{red}{\bf $\heart$} \& \textcolor{orange}{\bf RT} \& \textcolor{orange}{\bf @coldplay} \& : \& Fix \& You \& from \& the \& back \& \textcolor{gray}{\bf \#ColdplayMinneapolis} \& \textcolor{blue}{\bf http://bit.ly/2dj2WCl} \& Nice \& \textcolor{cyan}{\bf p} \& ...\\
			\tiny ADJ \&\tiny SYM \&\tiny X \&\tiny X \&\tiny PUNCT \&\tiny VERB \&\tiny NOUN \&\tiny ADP \&\tiny DET \&\tiny NOUN \&\tiny X \&\tiny X \& \tiny ADJ \& \tiny X \& \tiny PUNCT \\
		\end{deptext}
		\deproot{1}{root}
		\depedge[edge unit distance=1em]{1}{2}{discourse}
		\deproot{6}{root}
		\depedge[edge unit distance=1em]{6}{3}{discourse}
		\depedge[edge unit distance=1em]{3}{4}{discourse}
		\depedge[edge unit distance=1em]{3}{5}{punct}
		\depedge[edge unit distance=1em]{6}{7}{obj}
		\depedge[edge unit distance=0.8em]{6}{10}{obl}
		\depedge[edge unit distance=1em]{10}{8}{case}
		\depedge[edge unit distance=1em]{10}{9}{det}
		\depedge[edge unit distance=0.8em]{6}{11}{discourse}
		\depedge[edge unit distance=0.8em]{6}{12}{list}
		\deproot{13}{root}		
		\depedge[edge unit distance=0.8em]{13}{14}{discourse}
		\depedge[edge unit distance=0.8em]{13}{15}{punct}
	\end{dependency}
	\caption{An example to illustrate non-syntactic tokens:
		\textcolor{red}{sentiment emoticon},
		\textcolor{orange}{retweet marker and its following at-mention},
		\textcolor{gray}{topical hashtag},
		\textcolor{blue}{referential URL}, and
		\textcolor{cyan}{truncated word}.
		This is a concatenation of three real tweets.
	}\label{fig:non-syn-toks}
\end{figure*}

\begin{figure*}[t]
	\centering
	\small
	\begin{dependency}[edge slant=2, text only label, label style=above]
		\begin{deptext}
			@username \& its \& \#awesome \& u \& gonna \& $\heart$ \& it \& Chk \& out \& our \& cooool \& project \& on \& http://project\_link \& + \& RT \& it\\
			\tiny PROPN \& \tiny PRON \& \tiny ADJ \& \tiny PRON \& \tiny VERB \& \tiny VERB \& \tiny PRON \& \tiny VERB \& \tiny ADP \& \tiny PRON \& \tiny ADJ \& \tiny NOUN \& \tiny ADP \& \tiny X \& \tiny SYM \& \tiny VERB\& \tiny PRON\\
		\end{deptext}
		\deproot{3}{root}
		\depedge[edge unit distance=1em]{3}{1}{vocative}
		\depedge[edge unit distance=1em]{3}{2}{nsubj}
		\depedge[edge unit distance=1em]{3}{5}{parataxis}
		\depedge[edge unit distance=1em]{5}{4}{nsubj}
		\depedge[edge unit distance=1em]{5}{6}{xcomp}
		\depedge[edge unit distance=1em]{6}{7}{obj}
		\depedge[edge unit distance=0.7em]{3}{8}{parataxis}
		\depedge[edge unit distance=1em]{8}{9}{compound:prt}
		\depedge[edge unit distance=0.7em]{8}{12}{obj}
		\depedge[edge unit distance=1em]{12}{10}{nmod:poss}
		\depedge[edge unit distance=1em]{12}{11}{amod}
		\depedge[edge unit distance=0.6em]{8}{14}{obl}
		\depedge[edge unit distance=1em]{14}{13}{case}
		\depedge[edge unit distance=0.55em]{8}{16}{conj}
		\depedge[edge unit distance=1em]{16}{15}{cc}
		\depedge[edge unit distance=1em]{16}{17}{obj}
	\end{dependency}
	\caption{An example to illustrate informal but syntactic tokens.
		This is a contrived example inspired by several tweets.}\label{fig:informal-toks}
\end{figure*}

@@

\begin{itemize}
%\item For the at-mentions in retweet constructions, we attach them to their retweet markers. We will discuss our choice in the following paragraph;
\item If a non-syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate. Retweet construction is a special case and we will discuss our choice in the following paragraph.
\item If the whole sentence is made of a sequence of non-syntactic tokens, we attach all these tokens to the first one.
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item Non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English dataset.% \yicomment{need to add punct for punctuations?}
%\\ \yicomment{For retweet at-mention case, we attach the retweet at-mentions to the RT sign, so that the retweet action and the people whose tweet is being retweeted could be connected.\footnote{We can also follow the first convention to attach the retweet at-mention to its predicate, but we think it is better for downstream tasks.}}
\end{itemize}

@@

\paragraph{Retweet construction.} Figure \ref{fig:non-syn-toks} shows
an example of the retweet construction (\emph{RT @coldplay :}).  This
might be treated as a verb phrase, with \emph{RT} as a verb and the
at-mention as an argument \yicomment{@coldplay seems not to be an
  object, could be some modifier like vocative, or some prepositional
  case (the tweets on/about coldplay)}\yjcomment{i am okey with
  current version} \nascomment{changed ``object'' to ``argument'' - better?}.
% \yicomment{or might be treated as 2 non-syntactic tokens and a punctuation, which should all attach to the predicate of the following retweeted tweet.}
%\yjcomment{i wouldn't like the readers to think in this direction.}.
  This solution would lead to an uninformative root
word and, since this expression is idiomatic to Twitter, might create
unnecessary confusion for downstream applications aiming to identify
the main predicate(s) of a tweet.  We therefore treat the whole
expression as non-syntactic, including assigning the \emph{other}
(\texttt{X}) part of speech to both \emph{RT} and \emph{@coldplay},
attaching the at-mention to \emph{RT} with the \emph{discourse}
label and the colon to \emph{RT} with the \emph{punct}(uation) label,
and attaching \emph{RT} to the predicate of the following sentence.
%\yicomment{since we have introduced retweet at-mentions in the non-syntactic tokens, here we can just briefly justify our choice.}.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & $F_1$ \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\		 
% 		UD pipe v1.2 & 94.7 \\
		\hdashline
		UD pipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}
\citet{wang-EtAl:2017:Long6} show that the parser trained 
on the combination of UD\_English and English dialect (Singlish) treebank
outperform the one only on the dialect treebank.  Preliminary results
showed the same in our case \nascomment{added to address an open issue}, so 
we follow their work and trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English training set
and tested on the \textsc{Tweebank v2} test set.% \yicomment{should mention that combination of corpora improves the results, otherwise we should answer why we didn't try training ONLY on Tweebank v2}\yjcomment{add discussion}
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540} systems are rule-based systems and were not adapted
to the UD tokenization scheme.
The UD pipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. Compared with the UD pipe, we use LSTM
instead of GRU in our model and we also use a larger size of hidden units (64 against 20),
which has stronger representation power. 
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results indicate the value of statistical modeling
in tokenization for informal texts.

@@

Part-of-speech tagging for tweets has been extensively studied \cite{ritter-EtAl:2011:EMNLP,gimpel-EtAl:2011:ACL-HLT2011,derczynski-EtAl:2013:RANLP-2013,owoputi-EtAl:2013:NAACL-HLT,gui-EtAl:2017:EMNLP20172}. We therefore consider existing POS taggers for tweets instead of developing our own.
On the annotation scheme designed in \S\ref{sec:pos-anno}, based on UD and adapted for
Twitter, we compared several existing systems: the Stanford CoreNLP tagger, 
\citet{owoputi-EtAl:2013:NAACL-HLT}'s word cluster enhanced tagger
(both greedy and CRF variants), and 
\citet{ma-hovy:2016:P16-1}'s neural network tagger 
which achieves the state-of-the-art performance on PTB.
\citet{gui-EtAl:2017:EMNLP20172} presented a state-of-the-art neural tagger for Twitter,
but their implementation works only with the PTB tagset
\nascomment{check here -- seems surprising to me, hopefully that's
  enough for reviewers}, so we exclude
it. % Thus, we didn't include their tagger in our comparison.\yjcomment{i put some words on \cite{gui-EtAl:2017:EMNLP20172} in case the reviewer challenge our neural/non-neural comparsion.}
 %and also two bi-LSTM systems similar
%to \citet{DBLP:journals/corr/HuangXY15} and \citet{lample-EtAl:2016:N16-1}, one using
%word vectors and the other using word and character vectors (see Figure \ref{fig:pos-model}).
All compared systems were re-trained on the combination of the UD\_English and 
\textsc{Tweebank v2} training sets. We use Twitter-specific GloVe embeddings released by
\citet{pennington-socher-manning:2014:EMNLP2014} in all neural taggers
and parsers.\footnote{\url{http://nlp.stanford.edu/data/glove.twitter.27B.zip}}
